Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)

### 2021-09-23 Prelim/prelim.Rmd
updated prelim slides

\begin{columns}
\begin{column}{0.5\textwidth}

\begin{itemize}
\item Record linkage is the task of identifying duplicate records over noisy datasets.

\item Easy with unique identifiers, difficult when faced with errors

\item Far ranging applications in business, public health, and human rights

\end{itemize}

\end{column}
\begin{column}{0.5\textwidth}  %%<--- here
    \begin{center}
     picture
     \end{center}
\end{column}
\end{columns}

@@

```{r out.width="100%"}
knitr::include_graphics("fs_graphic.png")
```

@@

# m's and u's

# Previous work

@@

- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records. When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p$.

@@

- Gibbs sampler over each record pair has complexity $O(n_A \times n_B \times F)$, but sampler over agreement pattern has complexity $O(P \times n_B \times F)$

- Speeds up posterior calculations and sampling

@@

- Hashing plus SEI can reduce memory requirements by >99%.
  - Simulation of $20,000 \times 20,000$ linkage problem with 4 fields. Naive approach requires 6.4GB of storage, hashing and SEI requires 90MB.

@@

# El Salvador

# Distributed computation

# Storage Efficient Indexing

>>>>

<!-- \begin{columns} -->
<!-- \begin{column}{0.5\textwidth} -->

<!-- \begin{itemize} -->
<!-- \item Record linkage is the task of identifying duplicate records over noisy datasets. -->

<!-- \item Easy with unique identifiers, difficult when faced with errors -->

<!-- \item Far ranging applications in business, public health, and human rights -->

<!-- \end{itemize} -->

<!-- \end{column} -->
<!-- \begin{column}{0.5\textwidth}  %%<--- here -->
<!--     \begin{center} -->
<!--      picture -->
<!--      \end{center} -->
<!-- \end{column} -->
<!-- \end{columns} -->

- Record linkage is the task of identifying duplicate records over noisy datasets.

- Easy with unique identifiers, difficult when faced with errors

- Far ranging applications in business, public health, and human rights

# Examples

```{r}
knitr::include_graphics("rl_examples.png")
```

@@

```{r out.width="100%"}
knitr::include_graphics("fs_graphic.png")
```
# Previous work

@@

@@

- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns, regardless of number of records. When $(i,j)$ pair exhibits agreement pattern $p$, say $h(i, j)=p.$

<!-- (i, j) \in h_p$. -->

@@

- Run Gibbs sampler at level of agreement patterns, not record pairs
  - Sample the agreement of pattern $h(z_j, j)$, instead of record label $z_j$.
  - Use number of matches for each pattern to update $m$ and $u$
  - Back fill record labels at the end through $r_{p_j}$
  - Computational complexity $O(P \times n_B \times F)$ instead of $O(n_A \times n_B \times F)$

<!-- - Gibbs sampler over each record pair has complexity $O(n_A \times n_B \times F)$, but sampler over agreement pattern has complexity $O(P \times n_B \times F)$ -->

@@

- Hashing plus SEI can reduce memory requirements by >99%.
  - Simulation of $20,000 \times 20,000$ linkage task with 4 fields. Naive approach requires 6.4GB of storage for all-to-all comparisons, hashing and SEI requires 90MB.

# Three Simulations

- We compare \texttt{fabl} against \texttt{BRL} in three simulation studies
  - Measure precision and recall on 100 simulated datasets and varying levels of error and duplication across files
  - Measure speed when both $n_A$ and $n_B$ are increasing
  - Measure speed when $n_A$ is increasing and $n_B = 500$ is fixed.

@@

- Details of intuition behind $m$ and $u$ parameters

- Loss function and Bayes estimate

- Details of one-to-one matching; conflict resolution under \texttt{fabl};

- El Salvador case study (time under different values of $P$, violations of one-to-one matching)

- Further details on distributed computation of comparison vectors and storage efficient indexing
